Backscatter Modulation Design for Symbiotic Radio Networks

### 2021-09-02 main.tex
:pencil: Add remarks on decoding order

								to[short] ++(-0.75,0)
								to[short] ++(0,0.75) node[bareantenna](A){};

@@

							\draw (WI.north east) ++(0.25,0) node[]{$\vec{E}_{\mathrm{I}}$};
							\draw (WI.south east) ++(0.25,0) node[]{$\vec{H}_{\mathrm{I}}$};

@@

							\draw (WR.north east) ++(0.25,0) node[]{$\vec{E}_i$};
							\draw (WR.south east) ++(0.25,0) node[]{$\vec{H}_i$};

@@

								to [R=$Z_{\mathrm{A}}$] ++(0,-2)
								to [R=$Z_i$] ++(2,0) node[ground]{};

@@

			Consider a bistatic backscatter system that consists of an excitation source, a dedicated reader, and a passive tag. The excitation source generates a carrier wave signal, the dedicated reader decodes the tag message, and the tag simultaneously harvests energy, backscatters its own message, and demodulates the downlink information if necessary. As shown in Fig.~\subref*{fi:block_diagram}, a typical passive tag consists of a scattering antenna, an energy harvester, a integrated receiver\footnote{For example, \cite{Kim2021a} prototyped a compact-size pulse position demodulator based on an envelope detector, which brings great potential to coordination, synchronization, and reflection pattern control.}, a load-switching modulator, and on-chip components (e.g., micro-controller, memory, and sensors). A portion of the impinging signal is absorbed by the tag while the remaining is backscattered to the space, as illustrated in Fig.~\subref*{fi:one_port_network_model}. According to Green's decomposition \cite{Hansen1989}, the backscattered signal can be decomposed into the \emph{structural mode} component and the \emph{antenna mode} component. The former is fixed and depends on the antenna geometry and material properties\footnote{We assume the structural mode reflection can be modeled as part of the environment multipath and covered by channel estimation \cite{Boyer2014}.}, while the latter is adjustable and depends on the mismatch of the antenna and load impedance. Hence, the equivalent reflection coefficient at tag state $i$ is defined as\footnote{We assume the linear backscatter where $\Gamma_i$ is irrelevant to the incident electromagnetic field at the tag \cite{Dobkin2012}.}

@@

	\begin{section}{System Model}
			\def\svgwidth{0.9\columnwidth}
			\import{assets/}{symbiotic_radio.pdf_tex}
			\caption{A single-user single-tag symbiotic radio system.}
			\label{fi:symbiotic_radio}
		% As shown in Fig.~\ref{fi:system}, we propose a single-user (UE) single-tag (TG) symbiotic radio network where the RF signal generated by a $Q$-antenna Access Point (AP) is shared by two coexisting systems. In the primary AP-UE downlink system, the AP broadcasts to the single-antenna user. In the secondary AP-TG-UE backscatter system, the user serves as the reader of $L$ nearby single-antenna tags, while tag $l_k \in \mathcal{L}_k \triangleq \{1,\ldots,L_k\}$ harvests energy from and modulates information over the ambient RF signal by varying its reflection coefficient $\Gamma_{l_k}$\footnote{Due to the product-distance path loss model corresponding to finite-size tag reflection \cite{Ozdogan2020}, the backscatter dyadic channel deteriorates fast when the forward/backward distance increases or the number of hops increases \cite{Griffin2008}. Therefore, we assume each tag only contributes to its nearest user (i.e., creates no interference to other users), and omit all signals backscattered by two or more times.}. Denote the AP-UE $k$ direct channel as $\boldsymbol{h}_{\mathrm{D},k}^H \in \mathbb{C}^{1 \times Q}$, the AP-TG $l_k$ forward channel as $\boldsymbol{h}_{\mathrm{F},{l_k}}^H \in \mathbb{C}^{1 \times Q}$, and the TG $l_k$-UE $k$ backward channel as $h_{\mathrm{B},l_k} \in \mathbb{C}$. We consider the quasi-static block fading model and assume the CSI of all direct, forward and backward channels are known at the AP.\footnote{Due to the lack of RF chains at the passive tags, accurate and efficient estimation for the forward and backward channels can be challenging. Our assumption is based on the following reasoning. First, the direct AP-UE channel and the dyadic AP-TG-UE channel can be estimated with the AP sending pilots, the tags responding in a pre-defined frame-based manner, and the users performing least-square estimation within each frame\cite{Bharadia2015,Yang2015b,Guo2019e}. Second, the UE-TG-UE channel can be estimated similarly with the full-duplex users sending pilots\cite{Mishra2019}. Third, the users seek reciprocity to recover the TG-UE backward channel from the estimated UE-TG-UE channel\cite{Mishra2019b,Zawawi2019}, then extracts the AP-TG forward channel from the estimated AP-TG-UE channel provided by the AP. The whole process can be accelerated by the parallel estimation protocol proposed in \cite{Jin2021}.} Define $s_k$ as the primary symbol for user $k$ with $\boldsymbol{s} \triangleq [s_1,\ldots,s_K]^H \in \mathbb{C}^{K \times 1}$ satisfying $\mathbb{E}\{\boldsymbol{s}\boldsymbol{s}^H\}=\boldsymbol{I}$, and $c_{l_k}$ as the secondary symbol of tag $l_k$ with $\boldsymbol{c}_k \triangleq [c_1,\ldots,c_{L_k}]^H \in \mathbb{C}^{L_k \times 1}$ satisfying $\mathbb{E}\{\boldsymbol{c}_k\boldsymbol{c}_k^H\}=\boldsymbol{I}$. Let $s_k \sim \mathcal{CN} (0,1)$ and $c_{l_k}$ employs $M$-PSK by \eqref{eq:c}. As tags typically transmit at a much lower rate than the AP, we assume the secondary symbol period $T_c$ is $N \in \mathbb{Z}_{++}$ times the primary symbol period $T_s$ (i.e., $T_c=NT_s$) and focus on one particular secondary symbol period. For linear precoding with no dedicated tag beams, the transmit signal during the primary symbol period $n \in \mathcal{N} \triangleq \{1,\ldots,N\}$ is

>>>>

								to[short] ++(-1,0)
								to[short] ++(0,0.75) node[bareantenna](A){Ant};

@@

							\draw (WI.north east) ++(0.25,0) node[font=\Large]{$\vec{E}_{\mathrm{I}}$};
							\draw (WI.south east) ++(0.25,0) node[font=\Large]{$\vec{H}_{\mathrm{I}}$};

@@

							\draw (WR.north east) ++(0.25,0) node[font=\Large]{$\vec{E}_i$};
							\draw (WR.south east) ++(0.25,0) node[font=\Large]{$\vec{H}_i$};

@@

								to [R=$Z_{\mathrm{A}}$,font=\Large,/tikz/circuitikz/bipoles/length=1cm] ++(0,-2)
								to [R=$Z_i$,font=\Large,/tikz/circuitikz/bipoles/length=1cm] ++(2,0) node[ground]{};

@@

			Consider a bistatic backscatter system that consists of an excitation source, a dedicated reader, and a passive tag. The excitation source generates a carrier wave signal, the dedicated reader decodes the tag message, and the tag simultaneously harvests energy, backscatters its own message, and demodulates the downlink information if necessary. As shown in Fig.~\subref*{fi:block_diagram}, a typical passive tag consists of a scattering antenna, an energy harvester, a integrated receiver\footnote{For example, \cite{Kim2021a} prototyped a compact-size pulse position demodulator based on an envelope detector, which brings great potential to coordination, synchronization, and reflection pattern control.}, a load-switching modulator, and on-chip components (e.g., micro-controller, memory, and sensors). A portion of the impinging signal is absorbed by the tag while the remaining is backscattered to the space, as illustrated in Fig.~\subref*{fi:one_port_network_model}. According to Green's decomposition \cite{Hansen1989}, the backscattered signal can be decomposed into the \emph{structural mode} component and the \emph{antenna mode} component. The former is fixed and depends on the antenna geometry and material properties\footnote{We assume the structural mode reflection can be modeled as part of the environment multipath and covered by channel estimation \cite{Boyer2014}.}, while the latter is adjustable and depends on the mismatch of the antenna and load impedance. Hence, the equivalent reflection coefficient at tag state $i$ is defined as\footnote{We assume the linear backscatter model where $\Gamma_i$ is irrelevant to the incident electromagnetic field at the tag \cite{Dobkin2012}.}

@@

	\begin{section}{System Model}
		\begin{figure}[!t]
			\centering
			\def\svgwidth{0.9\columnwidth}
			\import{assets/}{symbiotic_radio.pdf_tex}
			\caption{A single-user single-tag symbiotic radio system.}
			\label{fi:symbiotic_radio}
		\end{figure}
		As shown in Fig.~\ref{fi:symbiotic_radio}, we propose a single-user (UE) single-tag (TG) symbiotic radio network where the RF signal generated by the single-antenna Access Point (AP) is shared by two coexisting systems. In the primary AP-UE downlink system, the AP transmits to the single-antenna user. In the secondary AP-TG-UE backscatter system, the AP acts as the carrier emitter, the user serves as the backscatter reader, and the single-antenna tag modulates its information over the reradiated RF signal by varying the reflection coefficient. Denote the AP-UE direct channel as $h_{\mathrm{D}}$, the AP-TG forward channel as $h_{\mathrm{F}}$, and the TG-UE backward channel as $h_{\mathrm{B}}$. We consider the quasi-static block fading model and assume the CSI of the direct channel and the cascaded forward-backward channel $h_{\mathrm{C}} \triangleq h_{\mathrm{B}} h_{\mathrm{F}}$ are known at the AP.\footnote{Due to the lack of RF chains at the passive tag, accurate and efficient CSI acquisition at the AP can be challenging. One possible approach is that the AP sends known pilots, the tag responds in a pre-defined manner, and the user performs least-square estimation then feeds back to the AP \cite{Bharadia2015,Yang2015b,Guo2019e}.} It is assumed that the primary symbol $s$ follows standard CSCG distribution $\mathcal{CN}(0,1)$ and the secondary symbol $c$ employs $M$-PSK modulation by \eqref{eq:backscatter_modulation}. Let $P$ be the average transmit power of the AP. The user simultaneously captures the signal from both primary and secondary links as\footnote{We assume the time difference of arrival from the AP-UE path and the AP-TG-UE path are negligible compared to the symbol period \cite{Guo2019b,Liang2020,Long2020a}.}
		\begin{equation}
			y = \sqrt{P} h_{\mathrm{D}} s + \sqrt{\alpha P} h_{\mathrm{C}} s c + n,
		\end{equation}
		where $n \sim \mathcal{CN}(0,1)$ is the additive white Gaussian noise.

		\begin{remark}
			The symbiotic radio network can be regarded as as a special case of Multiple Access Channel (MAC) because the AP and the tag transmit to the user simultaneously. It is known that SC-SIC with different decoding orders can achieve different vertices of the MAC capacity region \cite{Goldsmith2005}. Therefore, most relevant papers proposed to decode the primary message first (by treating the tag interference as noise), cancel out the primary contribution from the received signal, then decode the secondary message. Since the direct channel is typically much stronger than the cascaded channel \cite{Ozdogan2020}, the primary decoding should enjoy a high Signal-to-Interference-and-Noise Ratio (SINR) and the secondary decoding should be interference-free in the ideal case.
		\end{remark}

		\begin{remark}
			To investigate how backscatter modulation potentially benefits the primary transmission, we instead decode the secondary link before the primary link. Once the tag message is successfully decoded, the uncertainty of the AP-TG-UE path can be perfectly removed (by combining the backscattered symbol and the cascaded channel), and its contribution can be modeled within the equivalent AP-UE channel. This is reminiscent of IRS-aided point-to-point transmission with CSI uncertainty (by backscatter modulation instead of channel estimation), where the passive tag simultaneously embeds its own message (embodied in secondary decoding) and assists the legacy transmission (embodied in primary decoding).

			[Channel over channel, randomness over randomness]
		\end{remark}

		% Define $s_k$ as the primary symbol for user $k$ with $\boldsymbol{s} \triangleq [s_1,\ldots,s_K]^H \in \mathbb{C}^{K \times 1}$ satisfying $\mathbb{E}\{\boldsymbol{s}\boldsymbol{s}^H\}=\boldsymbol{I}$, and $c_{l_k}$ as the secondary symbol of tag $l_k$ with $\boldsymbol{c}_k \triangleq [c_1,\ldots,c_{L_k}]^H \in \mathbb{C}^{L_k \times 1}$ satisfying $\mathbb{E}\{\boldsymbol{c}_k\boldsymbol{c}_k^H\}=\boldsymbol{I}$. Let $s_k \sim \mathcal{CN} (0,1)$ and $c_{l_k}$ employs $M$-PSK by \eqref{eq:reflection_coefficient}. As tags typically transmit at a much lower rate than the AP, we assume the secondary symbol period $T_c$ is $N \in \mathbb{Z}_{++}$ times the primary symbol period $T_s$ (i.e., $T_c=NT_s$) and focus on one particular secondary symbol period. For linear precoding with no dedicated tag beams, the transmit signal during the primary symbol period $n \in \mathcal{N} \triangleq \{1,\ldots,N\}$ is
